Fixing Translation Divergences in Parallel Corpora for Neural MT

### 2018-08-22 EMNLP/divergence.tex
Restent deux lignes en trop

\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Jean Senellart$^\dag$ ,\ \ Fran\c cois Yvon$^\ddag$\\
  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
  {\tt firstname.lastname@systrangroup.com}\\
  $^\ddag$LIMSI-CNRS / Universit\'e Paris-Saclay 91405 Orsay, France\\
  {\tt firstname.lastname@limsi.fr}}

@@

\begin{document}
\maketitle
\begin{abstract}

%http://www.airccj.org/CSCP/vol4/csit42503.pdf

Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
%In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
Such resource is not naturally existing, and because of the process necessary to compile a parallel corpus, it may contain multiple sentence pairs that are often not as parallel as one might assume.
This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
We present a neural network to predict sentence similarity that minimises a loss function based on word divergence.
We show that accurate predictions are obtained allowing divergent sentences to be filtered out.
Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
We evaluate the presented method on  English-French and  English-German machine translation tasks.
Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.

@@

Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems. 
In the case of Neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
Therefore, the quality of an MT engine is heavily dependent upon the amount but also the quality of parallel sentences.\footnote{Note that  recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with the need of parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT systems. However, these systems still lag far behind systems trained in a supervised fashion, as considered in this work.}

Unfortunately, parallel texts are scarce resources. 
There are relatively few language pairs for which parallel corpora of large sizes are available, and even for those pairs, the corpora come mostly from few restricted domains. 
To alleviate the lack of parallel data, several approaches have been developed in the last years. 
They range from methods using non-parallel, or comparable data 
~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,GregoireL18,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences. 
Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.

\begin{table}[ht]
\small
\center
\begin{tabular}{ c|l }
  \hline  
  \texttt{en} & \it{What do you feel}\bf{, Spock}\it{?} \\
  \texttt{fr} & \it{Que ressentez-vous?} \\
  \texttt{gl} & {\small \it{What do you feel?}} \\
  \hline
  \texttt{en} & \it{How much do you get paid?} \\
  \texttt{fr} & \it{T'es pay\'e combien} \bf{de l'heure}\it{?} \\
  \texttt{gl} & {\small \it{How much do you get paid per hour?}} \\
  \hline  
  \texttt{en} &  \bf{That seems a lot.} \\
  \texttt{fr} & \bf{40 livres?} \\
  \texttt{gl} & {\small \it{40 pounds?}} \\
  \hline  
  \texttt{en} & \it{I brought you} \bf{french fries}\it{!} \\
  \texttt{fr} & \it{Je t'ai rapport\'e des} \bf{saucisses}\it{!} \\
  \texttt{gl} & {\small \it{I brought you sausage!}} \\
  \hline
\end{tabular}
\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}). }
\label{tab:examples}
\end{table}

Table~\ref{tab:examples} gives some examples of English-French parallel sentences that are not completely semantically equivalent, extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
Divergences are outlined using bold letters. 
Different types of translation divergences exist in a parallel corpus:
Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
Some translations may be completely uncorrelated (third row);
Inaccurate translations also exist (fourth row). 
Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.

In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
The architecture of our network allows us to distinguish between different types of divergences which are common to many existing bi-texts.
The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.

The remainder of this paper is structured as follows. 
Section~\ref{related} overviews related work. 
We describe in detail the core of the neural divergence classifier in Section~\ref{sec:divergence}. 
We report experiments with the presented model in Section~\ref{experiments}.
Section~\ref{sec:results} evaluates results. 
Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
All the code used in this paper is freely available\footnote{https://github.com/jmcrego/similarity}.

@@

Attempts to measure the impact of translation divergences in MT systems have focused on the introduction of noise in sentence alignments~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels. 
In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.

Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work. 
In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task. 
With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions. %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.

Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
The same authors have recently updated their work in~\cite{N18-1136}. 
The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. 
Our work differs from the previous as we make use of a network with a different topology. 
We model sentence similarity by means of optimising a loss function based on word alignments. 
Furthermore, the network predicts word similarity scores that we further use to correct divergent sentences.

@@

The architecture of our model is inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the model.

\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{network}
    \caption{Illustration of the model.} 
    \label{network}
\end{figure}

@@

In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. 
The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. 
Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.

The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence, 
$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
At this point, a measure of similarity between sentence pairs can be obtained by cosine similarity:
\begin{equation}
    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
    \label{cosine}
\end{equation}
%\noindent where two vectors (embeddings) with the same orientation have a cosine similarity of $1$, while two vectors with opposed orientation have a similarity of $-1$, independent of their magnitude.

Our model is optimised through maximising word alignment scores between words of both sentences using aggregation functions that summarise the alignment scores for each source/target word. 
Similar to~\cite{W16-2207} alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, and aggregation functions are defined:   
\begin{equation}
\begin{split}
    aggr_s(i,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{j=1}^{J} e^{r * S(i,j)}\right) \\
    aggr_t(j,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{i=1}^{I} e^{r * S(i,j)}\right)
\end{split}
\label{aggregation}
\end{equation}

The loss function is defined as:
%\vspace{-8mm}
\begin{equation}
\begin{split}
\mathcal{L}(src,tgt) = \ \ \ \ & \\
    \sum_{i=1}^I log(1+e&^{aggr_s(i,S) * sign_i}) \ \ +\\
 + \sum_{j=1}^J log(1+e&^{aggr_t(j,S) * sign_j})
\end{split}
\label{loss_wemb}
\end{equation}
%\vspace{-8mm}

@@

Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $sign_i$ and target $sign_j$ words are needed.
Different kind of examples are used to train our network:
\begin{itemize}

\item As positive examples we use paired sentences of a parallel corpus. All words in paired sentences are labelled as parallel ($-1$). %, $sign_i=-1$ and $sign_j=-1$. 
\item As negative examples we use random unpaired sentences. 
In this case, all words are labelled as divergent ($+1$). %, $sign_i=+1$ and $sign_j=+1$. 
Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create negative examples following the next two methods:

\item We replace random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
The rationale behind this method is to keep the new sentences as grammatical as possible. 
Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:

\begin{table}[h]
\center
\begin{tabular}{ll}
src: & { \small \texttt{What do you feel ?}} \\
tgt: & { \small \texttt{Que ressentez-vous ?}} \\
\end{tabular}
\end{table}

We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):

\begin{table}[h]
\center
\begin{tabular}{ll}
src: & { \small \texttt{What do {\bf we \ want} ?}} \\
$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
tgt: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

Divergent words are shown in bold. 
Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to '\texttt{you feel}' in the original sentence pair.

\item Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a second sentence at the beginning (or end) of the source (or target) sentence pair. 
Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
Given the original sentence pair previously shown, we may want to build the next negative example by inserting the sentence '\texttt{Not .}' at the end of the original source sentence:

@@

\end{itemize}

Replace and insert methods are applied on either side of the sentence pairs.
To avoid that negative examples are easily predicted by using the difference in length of sentences, we restrict negative examples to have a difference in length not exceeding a ratio of $2.0$ ($3.0$ for shortest sentences).

@@

We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words. % (as in the first and second examples of table~\ref{tab:examples}). 
Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides. 
Hence considering as parallel $s_u^v$ and $t_x^y$. 
To find optimal source ($u, v$) and target ($x, y$) indices that enclose parallel segments within the original sentence, we implement:
\begin{equation*}
\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
\end{equation*}

The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.

@@

We filter out divergences found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. 
We also use the very noisy English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. 
Both corpora present many potential divergences.
To evaluate English-French performance we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
English-German performance is evaluated on the publicly available newstest-2017 En-De test set~\cite{W17-4717}, corresponding to news stories selected from online sources.

In order to better assess the quality of our classifier when facing different word divergences we collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
$200$ paired sentences;
$100$ unpaired sentences;
$100$ sentences with replace examples; and
$100$ sentences with insert examples as detailed in Section~\ref{negative}.
All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation. %, basically splitting-off punctuation.

@@

After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
Vocabularies consist of the $50,000$ more frequent words.
Word embeddings are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}. 
Size of embeddings is $E_s=E_t=256$ cells. 
Both Bi-LSTM use $256$-dimensional hidden representations ($E=512$).
We use $r=1.0$. 
Network optimization is done using the SGD method along with gradient clipping~\cite{Pascanu:2013:DTR:3042817.3043083}. 
For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
We run $10$ epochs and start decaying at each epoch by $0.8$ when the loss on validation set increases. 
Divergence is computed following equation~\ref{cosine}. 
For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
The same number of examples are always generated for each type of example ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert). 
Alignments needed for {\texttt R}eplace and {\texttt I}nsert methods are performed using \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align}.

@@

In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned by joining both language sides.
Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
Hidden layer size is $512$. 
The inner layer of feed forward network is of size $2048$. 
Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
Batch size is set to $3072$.
%The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. We stop training after $30$ epochs.

@@

Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
Column \texttt{PURI} shows results over the entire test set, mixing all type of examples. 
As expected, the best accuracy is also obtained by the system trained on all type of  examples.

>>>>

\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Jean Senellart$^\dag$ ,\ \ Fran\c cois Yvon$^\ddag$\\
  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
  {\tt firstname.lastname@systrangroup.com}\\
  $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France\\
  {\tt firstname.lastname@limsi.fr}}

@@

\begin{document}
\maketitle

\begin{abstract}
%http://www.airccj.org/CSCP/vol4/csit42503.pdf

Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
%In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
Such resources are scarce, and because of the processes involved in their preparation, they may contain sentence pairs that are not as parallel as one would expect.
This paper describes an unsupervised method for detecting such translation divergences in parallel sentences. We rely on a  neural network trained
to predict sentence similarity that minimises a loss function based on word divergence. Its predictions are sufficiently accurate to effectively filtered out divergent translations.
Furthermore, word similarity scores predicted by the network are used to identify and fix some partial divergences, yielding additional parallel segments.
We evaluate these methods for  English-French and  English-German machine translation tasks: 
neural MT systems trained on the filtered/corrected corpora outperform the MT systems trained on the original data.

@@

Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems. 
In the case of neural MT, a large neural network is trained through maximising a proxy of translation performance on a parallel corpus. 
Therefore, the quality of MT engines is heavily dependent on the amount but also the quality of available parallel sentences.\footnote{Recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT. These systems however lag far behind supervised systems, as considered in this work.}

Parallel texts are unfortunately, scarce resources:
There are relatively few language pairs for which parallel corpora of large sizes exist, and even for those pairs, available corpora only concern few restricted domains. 
To alleviate the lack of parallel data, several approaches have been developed over the years. 
They range from methods using non-parallel, or comparable data 
~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,GregoireL18,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714}, using automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences. 
Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
This problem is mostly ignored in MT, where parallel sentences are considered to convey the exact same meaning; yet it seems particularly important for neural MT engines~\cite{chen2016adaptation}.

\begin{table}[ht]
\small
\center
\begin{tabular}{ c|l }
  \hline  
  \texttt{en} & \it{What do you feel}\bf{, Spock}\it{?} \\
  \texttt{fr} & \it{Que ressentez-vous?} \\
  \texttt{gl} & {\small \it{What do you feel?}} \\
  \hline
  \texttt{en} & \it{How much do you get paid?} \\
  \texttt{fr} & \it{T'es pay\'e combien} \bf{de l'heure}\it{?} \\
  \texttt{gl} & {\small \it{How much do you get paid per hour?}} \\
  \hline  
  \texttt{en} &  \bf{That seems a lot.} \\
  \texttt{fr} & \bf{40 livres?} \\
  \texttt{gl} & {\small \it{40 pounds?}} \\
  \hline  
  \texttt{en} & \it{I brought you} \bf{french fries}\it{!} \\
  \texttt{fr} & \it{Je t'ai rapport\'e des} \bf{saucisses}\it{!} \\
  \texttt{gl} & {\small \it{I brought you sausage!}} \\
  \hline
\end{tabular}
\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}). Divergences are in bold letters. }
\label{tab:examples}
\end{table}

Table~\ref{tab:examples} gives some examples of English-French parallel sentences that are not completely semantically equivalent, extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}.

Multiples types of translation divergences are found in parallel corpora:
Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation;
Some translations may be completely uncorrelated (third row);
Inaccurate translations also exist (fourth row). 
Note that divergent translations can be due various reasons~\cite{C14-1055}, the study of which is beyond the scope of this paper.

In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity,  % in the form of continuous vectors. 
relying on a neural architecture (see \textsection~\ref{sec:divergence}) that is able to identify several types of common cross-lingual divergences. % which are common to many existing bi-texts.
The resulting embeddings are then used to measure semantic equivalence between sentences.
To evaluate our method, we show in \textsection~\ref{experiments} that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
We also show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy. All the code used in this paper is freely available.\footnote{https://github.com/jmcrego/similarity}

% Useless for a short
% The remainder of this paper is structured as follows. 
% Section~\ref{related} overviews related work. 
% We describe in detail the core of the neural divergence classifier in Section. 
% We report experiments with the presented model in Section
% Section~\ref{sec:results} evaluates results. 
% Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.

@@

Attempts to measure the impact of translation divergences in MT have focused on the introduction of noise in sentence alignments~\cite{goute2012}, showing that statistical MT is highly robust to noise, and that performance only degrades seriously at very high noise levels.  In contrast, neural MTs seem to be more sensitive to noise~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.

Efforts devoted to characterising the degree of semantic equivalence between two snippets of texts in the same or different languages are presented~\cite{conf/semeval/AgirreBCDGMRW16}. % a workshop devoted to an objective similar to our work. 
In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
The authors show that a simple SVM classifier exploiting such sentence representations achieves state-of-the-art results in a textual entailment task. 
With the same objective, the system of~\citet{N16-1108} uses multiple convolutional layers and models pairwise word interactions. %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.

Our work is inspired by~\citet{W17-3209}, who train a SVM-based cross-lingual divergence detector using word alignments and sentence length features. 
Their work shows that an NMT system trained only on non-divergent sentences yields slightly better translation scores, while requiring less training time. 
A follow-up study by the same authors~\cite{N18-1136} achieves even better results, using the neural architecture of \citet{N16-1108}.
Our work differs from theirs as we make use of a network with a different, arguably simpler, topology. 
We model sentence similarity by means of optimising a loss function based on word alignments. 
Furthermore, the network predicts word similarity scores that we further use to correct divergent sentences.

@@

The architecture of our network is inspired by the work on word alignment of~\citet{W16-2207}, using however contextual, rather than fixed, word embeddings (see Figure~\ref{network}).
\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{network}
    \caption{Illustration of the model.} 
    \label{network}
\end{figure}

@@

It computes the similarity of any source-target sentence pair $(s,t)$, where $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. 
The model is composed of 2~bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. 
Since both $net_s$ and $net_t$ take the same form, we only describe the former network: 
it outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a vector encoding the $i^{th}$ source word as $h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
In addition, the last forward/backward hidden states (in dark grey on Figure~\ref{network}) are also concatenated  to represent whole sentences 
$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
The similarity between sentence pairs can then be obtained using eg.\ the cosine similarity:
\begin{equation}
    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
    \label{cosine}
\end{equation}
%\noindent where two vectors (embeddings) with the same orientation have a cosine similarity of $1$, while two vectors with opposed orientation have a similarity of $-1$, independent of their magnitude.

Our model is trained to maximize word alignment scores between words in both sentences, using aggregation functions that summarise the alignment scores for each source/target word. 
Similar to~\cite{W16-2207}, alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, further aggregated as follows:
\begin{equation}
\begin{split}
    aggr_s(i,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{j=1}^{J} e^{r * S(i,j)}\right) \\
    aggr_t(j,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{i=1}^{I} e^{r * S(i,j)}\right)
\end{split}
\label{aggregation}
\end{equation}

The training loss function is then defined as:
%\vspace{-8mm}
\begin{equation}
\begin{split}
\mathcal{L}(src,tgt) = \ \ \ \ & \\
    \sum_{i=1}^I log(1+e&^{aggr_s(i,S) * sign_i}) \ \ +\\
 + \sum_{j=1}^J log(1+e&^{aggr_t(j,S) * sign_j})
\end{split}
\label{loss_wemb}
\end{equation}
%\vspace{-8mm}

@@

Training is performed by minimizing Eq.~\eqref{loss_wemb}, for which annotated examples of source ($\operatorname{sign}_i$) and target ($\operatorname{sign_j}$) words are needed.
As positive examples, we use paired sentences of a parallel corpus; all words in such sentences are labelled as parallel ($ \forall i, j, \operatorname{sign}_i=\operatorname{sign}_j=-1$). %, $sign_i=-1$ and $sign_j=-1$. 
We consider three types of negative instances: the basic case uses random unpaired sentences; in this case, all words are labelled as divergent ($ \forall i, j, \operatorname{sign}_i=\operatorname{sign}_j=+1.$). %, $sign_i=+1$ and $sign_j=+1$. 
Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create more difficult negative examples as follows.

We first replace random sequences of words in source or target by a sequence of words with the same part-of-speeches.\footnote{The rationale is to try to keep the generated sentences as grammatical as possible; 
Otherwise, the network could learn to flag non-grammatical sentences as non-parallel.}
 Words that are not replaced are deemed parallel ($\operatorname{sign}_i=-1$) while those replaced are annotated as $\operatorname{sign}_i=+1$. 
Words aligned to some replaced words are also assigned the divergent label ($\operatorname{sign}_i=+1$). For instance, given the original sentence pair:

\begin{table}[h]
\center
\begin{tabular}{ll}
src: & { \small \texttt{What do you feel ?}} \\
tgt: & { \small \texttt{Que ressentez-vous ?}} \\
\end{tabular},
\end{table}

\noindent we may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'), yielding a new negative instance (divergent words are in bold):  
\begin{table}[h]
\center
\begin{tabular}{ll}
src: & { \small \texttt{What do {\bf we \ want} ?}} \\
$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
tgt: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

% Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
Note that we need word alignments to identify as divergent the sequence '\texttt{ressentez-vous}', which was aligned to '\texttt{you feel}' in the original sentence.
Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a sentence at the beginning (or end) of the source (or target) sentence. 
Words in the original sentence pair are annotated $\operatorname{sign}_i=-1$, while the new words inserted are considered divergent ($\operatorname{sign}_i=+1$).
Given the same sentence pair as above, a negative example is created by inserting the sentence '\texttt{Not .}' at the end of the original source:

@@

% Replace and insert methods are applied on either side of the sentence pairs. 
To finally avoid the generation of easy negative sentence pairs having a large difference in sentence length, we restrict negative examples to have a length ratio $< 2.0$ ($3.0$ for shortest sentences).

@@

Our training corpora contains many divergent sentences that follow a common pattern, consisting of adding some extra leading/trailing words. % (as in the first and second examples of table~\ref{tab:examples}). 
Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides. 
% Hence considering as parallel $s_u^v$ and $t_x^y$. 
To find optimal source ($u, v$) and target ($x, y$) indices that enclose parallel segments within the original sentence, we compute:
\begin{equation*}
\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
\end{equation*}
The $\mathcal{N}$-best sequences ($s_u^v$, $t_x^y$) are considered as likely corrections, in which we use the one having the highest similarity score to replace the original $(s_1^I, t_1^J)$.
Note that short sentences are not considered and we enforce $v - u > \tau$ and $y - x > \tau $. 
Figure~\ref{matrix} (left) displays an example of an alignment matrix $S(i,j)$.
An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
corresponding to $u=1$, $v=5$, $x=1$ and $y=3$.

@@

We filter out divergences from the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. 
We also use the very noisy English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. Both corpora present many potential divergences.
To evaluate English-French translation performance, we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual Skype conversations~\cite{mslt-corpus-iwslt-2016-release}. 
English-German performance is evaluated on the publicly available Newstest-2017~\cite{W17-4717}, corresponding to news stories selected from online sources.

In order to better assess the quality of our classifier when facing different word divergences, we also collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
$200$ paired sentences;
$100$ unpaired sentences;
$100$ sentences with replace examples; and
$100$ sentences with insert examples (see \textsection~\ref{negative}).
All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation. %, basically splitting-off punctuation.
After tokenisation, each out-of-vocabulary word is mapped to a special UNK token, assuming a vocabulary containing the $50,000$ more frequent words.

@@

Word embeddings of $E_s=E_t=256$ cells are initialised using \texttt{fastText},\footnote{https://github.com/facebookresearch/fastText} further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method of~\citet{lample2018word}. Both bi-LSTMs use $256$-dimensional hidden representations ($E=512$).
Network optimization is done using SGD with gradient clipping~\cite{Pascanu:2013:DTR:3042817.3043083}. 
For each epoch, we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
We run $10$ epochs and start decaying at each epoch by $0.8$ when the loss on validation set increases. 
Divergence is computed as in equation~\eqref{cosine} and setting $r=1.0$ ; 
For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
The same number of examples are always generated for each type of example ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert). 
Alignments needed for {\texttt R}eplace and {\texttt I}nsert methods are performed using \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align}.

@@

In addition to the basic tokenisation detailed above, we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned by joining both language sides.
Neural systems are based on the open-source project  \texttt{OpenNMT};  using a Transformer model similar to the model of~\citet{vaswani2017attention}: both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ head;
the hidden layer size is $512$; and the inner layer of feed forward network is of size $2048$. 
Word embeddings have $512$ cells. We set the dropout probability to $0.1$ and the batch size to $3072$.
%The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. Training stops after $30$ epochs.

@@

Models that were trained with the matching examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (in bold letters).
Column \texttt{PURI} gives results for the complete test set, mixing all type of examples. 
As expected, the best accuracy is also obtained when training on all types of  examples.
